---
title: "harinris_Homework1_lab"
author: "Harin Rishabh"
date: "2024-02-09"
output:
  html_document: default
  pdf_document: default
---

# 12.5 Lab: Unsupervised Learning
## 12.5.1 Principal Components Analysis

We are using the USArrests Dataset. We are calculating mean and variance.
```{r}
states <- row.names(USArrests)
apply(USArrests , 2, mean)
apply(USArrests , 2, var)
```

We now perform principal components analysis using the prcomp() function. pr.out$rotation contains the corresponding principal component
loading vector. Then we plot the first two principal components.
```{r}
pr.out <- prcomp(USArrests , scale = TRUE)
names(pr.out)
pr.out$rotation
biplot(pr.out , scale = 0)
```

Calculating STD and var of each principal component.
```{r}
pr.out$sdev
pr.var <- pr.out$sdev^2
pr.var
```

Computing the proportion of variance explained by each principal component. We see that the first principal component explains 62.0% of the variance in the data, the next principal component explains 24.7% of the variance, etc.
```{r}
pve <- pr.var / sum(pr.var)
pve
```

Plotting the PVE explained by each component, as well as the cumulative PVE.
```{w}
plot(pve , xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0, 1), type = "b")
plot(cumsum(pve), xlab = "Principal Component", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0, 1), type = "b")
```

## 12.5.2 Matrix Completion

Turning the data frame into a matrix, after centering and scaling each column to have mean zero and variance one.
```{r}
X <- data.matrix(scale(USArrests))
pcob <- prcomp(X)
summary(pcob)
```

Using SVD
```{r}
sX <- svd(X)
names(sX)
round(sX$v, 3)
pcob$rotation
```

We now omit 20 entries in the 50 × 2 data matrix at random. We do so by first selecting 20 rows (states) at random, and then selecting one of
the four entries in each row at random. This ensures that every row has at least three observed values.
```{r}
nomit <- 20
set.seed (15)
ina <- sample(seq (50) , nomit)
inb <- sample (1:4, nomit , replace = TRUE)
Xna <- X
index.na <- cbind(ina , inb)
Xna[index.na] <- NA
```

We first write a function that takes in a matrix, and returns an approximation to the matrix using the svd() function.
```{r}
fit.svd <- function(X, M = 1) {
  svdob <-svd(X)
  with(svdob ,
  u[, 1:M, drop = FALSE] %*%
  (d[1:M] * t(v[, 1:M, drop = FALSE ])))
}
```

Step 1 of algorithm
```{r}
Xhat <- Xna
xbar <- colMeans(Xna , na.rm = TRUE)
Xhat[index.na] <- xbar[inb]
```

Measuring progress of our iterations
```{r}
thresh <- 1e-7
rel_err <- 1
iter <- 0
ismiss <- is.na(Xna)
mssold <- mean (( scale(Xna , xbar , FALSE)[!ismiss])^2)
mss0 <- mean(Xna[!ismiss ]^2)
```

Step 2
```{r}
while(rel_err > thresh) {
  iter <- iter + 1
# Step 2(a)
  Xapp <- fit.svd(Xhat , M = 1)
# Step 2(b)
  Xhat [ ismiss ] <- Xapp [ ismiss ]
# Step 2(c)
  mss <- mean ((( Xna - Xapp)[! ismiss ])^2)
  rel_err <- ( mssold - mss ) / mss0
  mssold <- mss
  cat("Iter:", iter, "MSS:", mss,
  "Rel. Err:", rel_err, "\n")
}
```

Finally, we compute the correlation between the 20 imputed values and the actual values.
```{r}
cor(Xapp[ismiss], X[ismiss])
```

## 12.5.3 Clustering

We begin with a simple simulated example in which there truly are two clusters in the data: the first 25 observations have a mean shift relative to the next 25 observations.
```{r}
set.seed (2)
x <- matrix(rnorm (50 * 2), ncol = 2)
x[1:25, 1] <- x[1:25, 1] + 3
x[1:25, 2] <- x[1:25, 2] - 4
```

We now perform K-means clustering with K = 2.
```{r}
km.out <- kmeans(x, 2, nstart = 20)
km.out$cluster
par(mfrow = c(1, 2))
plot(x, col = (km.out$cluster + 1), main = "K-Means Clustering Results with K = 2", xlab = "", ylab = "", pch = 20, cex = 2)
```

Clustering with k=3.
```{r}
set.seed (4)
km.out <- kmeans(x, 3, nstart = 20)
km.out
plot(x, col = (km.out$cluster + 1),
main = "K-Means Clustering Results with K = 3",
xlab = "", ylab = "", pch = 20, cex = 2)
```

## Hierarchial Clusterig
```{r}
hc.complete <- hclust(dist(x), method = "complete")
hc.average <- hclust(dist(x), method = "average")
hc.single <- hclust(dist(x), method = "single")
par(mfrow = c(1, 3))
plot(hc.complete , main = "Complete Linkage",
xlab = "", sub = "", cex = .9)
plot(hc.average , main = "Average Linkage",
xlab = "", sub = "", cex = .9)
plot(hc.single, main = "Single Linkage",
xlab = "", sub = "", cex = .9)
```

To determine the cluster labels for each observation associated with a given cut of the dendrogram, we can use the cutree() function.
```{r}
cutree(hc.complete , 2)
cutree(hc.average , 2)
cutree(hc.single, 2)
cutree(hc.single, 4)
```

Scaling
```{r}
xsc <- scale(x)
plot(hclust(dist(xsc), method = "complete"),
main = "Hierarchical Clustering with Scaled Features")
```

Correlation-based distance can be computed using the as.dist() function, which converts an arbitrary square symmetric matrix into a form that
the hclust() function recognizes as a distance matrix.
```{r}
x <- matrix(rnorm (30 * 3), ncol = 3)
dd <- as.dist (1 - cor(t(x)))
plot(hclust(dd, method = "complete"),
main = "Complete Linkage with Correlation -Based Distance",
xlab = "", sub = "")
```